Fractional Types

### 2013-02-24 frac.tex
hasegawa

For these reasons, it is interesting to consider other possible semantic
interpretations of fractionals. A natural alternative is to use the
``canonical'' compact closed category, that is finite dimensional vector
spaces and linear
maps~\cite{Selinger:2011:FDH:1942319.1942398,Hasegawa:2008:FDV:1805839.1805859}
over fields of characteristic $0$ (or even the category of finite dimensional
Hilbert spaces). Let us fix an arbitrary field $k$ of characteristic $0$.
Then each type {{b}} in {{PiTF}} is interpreted as a finite dimensional
vector space $\mathbf{V}_b$ over the field $k$. In particular:
\begin{itemize}
  \item every vector space contains a zero vector which means that the type
    {{0}} (if included in {{PiTF}}) would not be the ``empty'' type.
    Furthermore all combinators would be
    strict as they would have to map the zero vector to itself. 
  \item the type {{1}} is interpreted as a 1-dimensional vector space and hence
    is isomorphic to the underlying field.
  \item the fractional type {{1/b}} is interpreted as the dual vector space
    to the vector space $\mathbf{V}_b$ representing {{b}} consisting of all
    the \emph{linear functionals} on $\mathbf{V}_b$.
  \item one can then validate certain desirable properties: {{1/(1/b)}} is
    isomorphic to~{{b}}; and {{eps*_b}} corresponds to a bilinear form which
    maps a dual vector and a vector to a field element.
\end{itemize}
In such categories, the fractional type is given a non-trivial
interpretation. Indeed, while the space of column vectors is isomorphic to
the space of row vectors, they are nevertheless quite different, being
$(1,0)$ and $(0,1)$ tensors (respectively).  In other words, the category
provides a more refined model in which isomorphic negative-information values
and positive-information are not identified.

>>>>

For these reasons, it is interesting to consider other possible semantic
interpretations of fractionals. A natural alternative is to use the
``canonical'' compact closed category, that is finite dimensional vector
spaces and linear
maps~\cite{Selinger:2011:FDH:1942319.1942398,hasegawa2008finite}
over fields of characteristic $0$ (or even the category of finite dimensional
Hilbert spaces). Let us fix an arbitrary field $k$ of characteristic $0$.
Then each type {{b}} in {{PiTF}} is interpreted as a finite dimensional
vector space $\mathbf{V}_b$ over the field $k$. In particular:
\begin{itemize}
  \item every vector space contains a zero vector which means that the type
    {{0}} (if included in {{PiTF}}) would not be the ``empty'' type.
    Furthermore all combinators would be
    strict as they would have to map the zero vector to itself. 
  \item the type {{1}} is interpreted as a 1-dimensional vector space and hence
    is isomorphic to the underlying field.
  \item the fractional type {{1/b}} is interpreted as the dual vector space
    to the vector space $\mathbf{V}_b$ representing {{b}} consisting of all
    the \emph{linear functionals} on $\mathbf{V}_b$.
  \item one can then validate certain desirable properties: {{1/(1/b)}} is
    isomorphic to~{{b}}; and {{eps*_b}} corresponds to a bilinear form which
    maps a dual vector and a vector to a field element.
\end{itemize}
In such categories, the fractional type is given a non-trivial
interpretation. Indeed, while the space of column vectors is isomorphic to
the space of row vectors, they are nevertheless quite different, being
$(1,0)$ and $(0,1)$ tensors (respectively).  In other words, the category
provides a more refined model in which isomorphic negative-information values
and positive-information are not identified.
